Learning Network Size while Training with ShrinkNets

### 2018-02-09 paper/ICML/appendix.tex
Resolve most of the bibliography

\end{proof}
\subsection{Experiment details}
\subsubsection{Multi-Target Linear and Logistic regressions}
\label{linear_datasets}
\subsubsection{Neuron Removal strategies}
\subsubsection{Convergence and Training Dynamics of Neural networks}
\subsubsection{Hyper-optimization of ShrinkNets}
\subsubsection{Performance}

>>>>

\end{proof}
\subsection{Additional Experiment}
As we showed, Group sparsity share similarities with our method, and we claim
that ShrinkNets are a relaxation of group sparsity.  In this experiment we want
to compare the two aproaches.  We decided to focus on multi-target linear
regression because in the single target case, groups in the Group Sparsity
problem would have a size of one ($\bm{A}$ would be a vector in this case).

The evaluation will be done on two datasets \texttt{scm1d} and \texttt{oes97}
\cite{Spyromitros-Xioufis2016} for linear regressions and we will use \texttt{gina\_prior2} \cite{4371065} and
the \textit{Gas Sensor Array Drift Dataset} \cite{VERGARA2012320} (that we shorten in
\texttt{gsadd}) for logistic regressions.

For each dataset we fit with different regularization parameters and measure
the error and sparsity obtained after convergence. In this context we define
sparsity as the ratio of columns that have all their weight under $10^{-3}$ in
absolute value. Regularization parameters were choosed in order to obtain the
widest sparsity spectrum. Loss is normalized depending on the problem to be in
the $[0, 1]$ range. We summarized the results in \cref{sparsity_accuracy}. From
our experiments it is clear that ShrinkNets can fit the data closer than Group
Sparsity for the same amount of sparsity. The fact that we are able to reach
very low loss demonstrate that even if our objective function is non convex, in
practice it works as good or better as convex alternatives.
\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{regressions}
\vspace*{-5mm}
\caption{\label{sparsity_accuracy}Loss/Sparsity trade off comparison between Group Sparsity and Shrinknet on linear and logistic regression. From top to bottom and left to right we show the results for \texttt{scm1d}, \texttt{oes97}, \texttt{gina\_prior2} and \texttt{gsadd}.}

\end{center}
\vspace*{-4mm}
\end{figure}

### 2018-02-09 paper/ICML/approach.tex
Resolve most of the bibliography

We have implemented \textbf{Switch Layers} and the associated training procedure
as a library in pytorch~\cite{github}.
The layer can be freely mixed with other popular layers such as convolutional
layers, batchnorm layers, fully connected layers etc.
There are two key challenges that must be overcome while implementing 
\textbf{Switch Layers}.
First, since ShrinkNets kill off a large fraction of neurons, we must remove the
overhead of carrying around dead neurons.
And second, we must update the various network layers and training machinery (e.g.
optimizer state) to reflect the deactivation of neurons.
We address these challenges by augmenting every layer with additional information
about neuron status and implementing a \emph{neural garbage collection} 
mechanism which prunes deactivated neurons on-the-fly and updates the state of 
the training machinery.
% We do so by augmenting every layer with additional information such as what
% neurons have been deactivated.
% The layer then updates its size accordingly and this change is then propagated
% to other parts of the training machinery including the optimizer which must
% know the sizes of parameters.

>>>>

We have implemented \textbf{Switch Layers} and the associated training procedure
as a library in pytorch~\cite{paszke2017automatic}. \gl{This is not a github citation but this is what they ask to cite}
The layer can be freely mixed with other popular layers such as convolutional
layers, batchnorm layers, fully connected layers etc.
There are two key challenges that must be overcome while implementing 
\textbf{Switch Layers}.
First, since ShrinkNets kill off a large fraction of neurons, we must remove the
overhead of carrying around dead neurons.
And second, we must update the various network layers and training machinery (e.g.
optimizer state) to reflect the deactivation of neurons.
We address these challenges by augmenting every layer with additional information
about neuron status and implementing a \emph{neural garbage collection} 
mechanism which prunes deactivated neurons on-the-fly and updates the state of 
the training machinery.
% We do so by augmenting every layer with additional information such as what
% neurons have been deactivated.
% The layer then updates its size accordingly and this change is then propagated
% to other parts of the training machinery including the optimizer which must
% know the sizes of parameters.

### 2018-02-09 paper/ICML/evaluation.tex
Resolve most of the bibliography

Previous attempt at solving this problem focus on simple datasets and simple
architectures (limited to Fully connected). We believe that self-resizing
network really matter for problems where the architecture you need to solve them
involve so many layers that it is impractical to explore the space of possible
sizes. We considered ImageNet but the time required to train models is so long
that it would have not been possible to train enough and get statistically
significant results. This is why we picked \texttt{CIFAR10} [ref]. It is an
image classification dataset containing $60000$ color images $(3 \times 32
\times 32)$, belonging to $10$ different classes.

To solve this task we use the \texttt{VGG16} model [ref]. It is constituted of
alternating convolutional layers and \textit{MaxPool} layers interleaved by
\textit{BatchNorm} [ref] and \textit{ReLU} [ref] layers. The two last layers are
Fully connected layers separated by just a \textit{ReLU} activation function.

@@

We assume no prior knowledge on the optimal batch size, learning rate,
$\lambda$ or weight decay ($\lambda_2$). This is why, we randomly sample them
from a range of reasonable values (\gl{should we make them explicit ?}).
Training is done using our library, based on \texttt{PyTorch} [ref], that
support dynamically resized layers. We train the network using gradient descent
and \textit{Adam} optimizer [ref]. We start with the learning rate sampled
randomly and every $5$ epochs of non improvement in validation accuracy we
divide the learning rate by $10$. We stop training after $400$ epochs or when
the learning rate is under $10^{-7}$ whichever comes first. \gl{Should I also
  give the details about the removal strategy we used and the $\gamma$ and
  threshold I used ? because this is clearly getting boring}
For each of the models we trained, we pick the epoch with the best validation
accuracy and report the corresponding testing accuracy. Because of the nature of
our method, it can happen that for network that are aggressively compressed, the
best validation accuracy is obtained early in the training, before the size has
converged. To be sure that accuracy measured corresponds to the final shape and
not the starting shape, we only consider the second half of the training when
picking the best epoch. For each model, we also measure the total size, in
number of floating point parameters, excluding the \textit{Switch Layers}
because as we saw in section \gl{REFERENCE\_ME}, we can get rid of them when
training is done.

@@

Our second dataset we will evaluate on is \texttt{COVERTYPE} [ref]. This dataset
contains $581012$ description of geographical area (elevation, inclination,
etc...) and the goal is to predict the type of forest growing there. We picked
this one for couple of reasons. First it is simple that we can reach good
accuracy with only a few layers. This important because we want to show that
\textit{ShrinkNets} find sizes as good as \textit{Static Networks}, even if we
are sampling the entire space of possible sizes. The second reason is that [ref]
also perform their evaluation on this dataset, which allow us to compare the
results obtained by the two methods.

The only difference in terms of setup are minimal. We use the same architecture
as [ref], i.e. three layers Fully Connected neural network with  no
\textit{Dropout} [ref] nor \textit{BatchNorm} [ref]. \gl{Should we say here that we
  don't expect Dropout to work here ? I could write an entire paragraph about it
  if needed}. In this case for the \textit{Static Networks}, we sample
independently sizes of the three different layers to explore all possible
architectures.

@@

We sucessfully demonstrated that the architecture \textit{Shrinknets} converge
to make sense both in term of size and performance. And for a given accuracy the
size needed is significantly smaller than when we use the classic heursitc we
commonly use to size convolutional neural networks. It might be an indication
that this strategy is suboptimal and might need to be adjusted. During our
experimentations on simpler datasets like \texttt{MNIST} and
\texttt{FashionMNIST} we observed that even if they have the same number of
classes, input features, and output distribution. For a fixed $\lambda$
\textit{ShrinkNets} converged to considerably bigger networks in the case of
\texttt{FashionMNIST}. This evidence shows that final architecture not only
depends on the output distribution or shape of the data but actually reflects
the dataset (we know that FASHIONMNIST is harder then MNIST)

@@

As we showed, Group sparsity share similarities with our method, and we claim
that ShrinkNets are a relaxation of group sparsity.  In this experiment we want
to compare the two aproaches.  We decided to focus on multi-target linear
regression because in the single target case, groups in the Group Sparsity
problem would have a size of one ($\bm{A}$ would be a vector in this case).

The evaluation will be done on two datasets \texttt{scm1d} and \texttt{oes97}
[ref] for linear regressions and we will use \texttt{gina\_prior2} [ref] and
the \textit{Gas Sensor Array Drift Dataset} [ref] (that we shorten in
\texttt{gsadd}) for logistic regressions.

For each dataset we fit with different regularization parameters and measure
the error and sparsity obtained after convergence. In this context we define
sparsity as the ratio of columns that have all their weight under $10^{-3}$ in
absolute value. Regularization parameters were choosed in order to obtain the
widest sparsity spectrum. Loss is normalized depending on the problem to be in
the $[0, 1]$ range. We summarized the results in \cref{sparsity_accuracy}. From
our experiments it is clear that ShrinkNets can fit the data closer than Group
Sparsity for the same amount of sparsity. The fact that we are able to reach
very low loss demonstrate that even if our objective function is non convex, in
practice it works as good or better as convex alternatives. Details about these
datasets and the parameters used are available in \cref{linear_datasets}.

\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{regressions}
\vspace*{-5mm}
\caption{\label{sparsity_accuracy}Loss/Sparsity trade off comparison between Group Sparsity and Shrinknet on linear and logistic regression. From top to bottom and left to right we show the results for \texttt{scm1d}, \texttt{oes97}, \texttt{gina\_prior2} and \texttt{gsadd}.}

>>>>

Previous attempt at solving this problem focus on simple datasets and simple
architectures (limited to Fully connected). We believe that self-resizing
network really matter for problems where the architecture you need to solve them
involve so many layers that it is impractical to explore the space of possible
sizes. We considered ImageNet but the time required to train models is so long
that it would have not been possible to train enough and get statistically
significant results. This is why we picked \texttt{CIFAR10} \cite{Krizhevsky2009}. It is an
image classification dataset containing $60000$ color images $(3 \times 32
\times 32)$, belonging to $10$ different classes.

To solve this task we use the \texttt{VGG16} model \cite{Srivastava2014}. It is constituted of
alternating convolutional layers and \textit{MaxPool} layers interleaved by
\textit{BatchNorm} [ref] and \textit{ReLU} [ref] layers. The two last layers are
Fully connected layers separated by just a \textit{ReLU} activation function.

@@

We assume no prior knowledge on the optimal batch size, learning rate,
$\lambda$ or weight decay ($\lambda_2$). This is why, we randomly sample them
from a range of reasonable values (\gl{should we make them explicit ?}).
Training is done using our library, based on \texttt{PyTorch} \cite{paszke2017automatic}, that
support dynamically resized layers. We train the network using gradient descent
and \textit{Adam} optimizer \cite{DBLP:journals/corr/KingmaB14}. We start with
the learning rate sampled randomly and every $5$ epochs of non improvement in
validation accuracy we divide the learning rate by $10$. We stop training after
$400$ epochs or when the learning rate is under $10^{-7}$ whichever comes
first. \gl{Should I also give the details about the removal strategy we used
  and the $\gamma$ and threshold I used ? because this is clearly getting
  boring} For each of the models we trained, we pick the epoch with the best
validation accuracy and report the corresponding testing accuracy. Because of
the nature of our method, it can happen that for network that are aggressively
compressed, the best validation accuracy is obtained early in the training,
before the size has converged. To be sure that accuracy measured corresponds to
the final shape and not the starting shape, we only consider the second half of
the training when picking the best epoch. For each model, we also measure the
total size, in number of floating point parameters, excluding the
\textit{Switch Layers} because as we saw in section \gl{REFERENCE\_ME}, we can
get rid of them when training is done.

@@

Our second dataset we will evaluate on is \texttt{COVERTYPE} \cite{Blackard:1998:CNN:928509}. This dataset
contains $581012$ description of geographical area (elevation, inclination,
etc...) and the goal is to predict the type of forest growing there. We picked
this one for couple of reasons. First it is simple that we can reach good
accuracy with only a few layers. This important because we want to show that
\textit{ShrinkNets} find sizes as good as \textit{Static Networks}, even if we
are sampling the entire space of possible sizes. The second reason is that [ref]
also perform their evaluation on this dataset, which allow us to compare the
results obtained by the two methods.

The only difference in terms of setup are minimal. We use the same architecture
as \cite{Scardapane2017}, i.e. three layers Fully Connected neural network with  no
\textit{Dropout} \cite{Srivastava2014} nor \textit{BatchNorm} \cite{DBLP:journals/corr/IoffeS15}. \gl{Should we say here that we
  don't expect Dropout to work here ? I could write an entire paragraph about it
  if needed}. In this case for the \textit{Static Networks}, we sample
independently sizes of the three different layers to explore all possible
architectures.

@@

We sucessfully demonstrated that the architecture \textit{Shrinknets} converge
to make sense both in term of size and performance. And for a given accuracy the
size needed is significantly smaller than when we use the classic heursitc we
commonly use to size convolutional neural networks. It might be an indication
that this strategy is suboptimal and might need to be adjusted. During our
experimentations on simpler datasets like \texttt{MNIST} \cite{Lecun1998} and
\texttt{FashionMNIST} \cite{Xiao2017} we observed that even if they have the same number of
classes, input features, and output distribution. For a fixed $\lambda$
\textit{ShrinkNets} converged to considerably bigger networks in the case of
\texttt{FashionMNIST}. This evidence shows that final architecture not only
depends on the output distribution or shape of the data but actually reflects
the dataset (we know that FASHIONMNIST is harder then MNIST)

@@

### 2018-02-09 paper/ICML/related_work.tex
Resolve most of the bibliography

The most popular techniques for hyperparameter optimization include simple 
methods such as random search~\cite{Bergstra2012} which have been shown to work
suprisingly well compared to more complex methods such as those based on
Bayesian optimization~\cite{Snoek12}.
Techniques such as~\cite{Snoek12} model generalization performance via 
Gaussian Processes~\cite{GaussianProcesses} and select hyperparameter 
combinations that come from uncertain areas of the hyperparameter space.
Recently, methods based on bandit algorithms (e.g. ~\cite{li2016hyperband, 
jamieson2016}) have also become a popular way to tune hyperparameters.
As noted before, all of the above techniques require many tens to hundreds of
models to be trained, making this process computationally inefficient and slow.

>>>>

The most popular techniques for hyperparameter optimization include simple 
methods such as random search~\cite{BergstraJAMESBERGSTRA2012} which have been shown to work
suprisingly well compared to more complex methods such as those based on
Bayesian optimization~\cite{Snoek12}.
Techniques such as~\cite{Snoek12} model generalization performance via 
Gaussian Processes~\cite{GaussianProcesses} and select hyperparameter 
combinations that come from uncertain areas of the hyperparameter space.
Recently, methods based on bandit algorithms (e.g. ~\cite{li2016hyperband, 
jamieson2016}) have also become a popular way to tune hyperparameters.
As noted before, all of the above techniques require many tens to hundreds of
models to be trained, making this process computationally inefficient and slow.
